---
title: "Test Slide Deck"
author: "Tyler J. Brough, Ph.D."
date: "March 28, 2018"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```


## Artificial Time Series Data

```{r, echo=TRUE}
N <- 100
e <- rnorm(100)
y <- rep(0, N)
rho <- 0.5
y[1] <- e[1] * sqrt(1 / (1 - rho^2))

for(t in 2:N)
{
    y[t] <- rho * y[t-1] + e[t]
}
```

## Slide with Longer Code Chunk

```{r, echo=TRUE}
m <- 10
q <- 1 / m
B <- 10000
N <- length(y)
ystar <- matrix(0, nrow=B, ncol=N)

for(b in 1:B) {
  u <- sample(1:N, size=N, replace=TRUE)
  for(t in 2:N) {
    v <- runif(1)
    if(v < q) { u[t] <- sample(1:N, size=1) }
    else {
      u[t] <- u[t-1] + 1
      if(u[t] > N) { u[t] <- u[t] - N }
    }
  }
  ystar[b,] <- y[u]
}

c(mean(y), mean(apply(ystar, 1, mean)))
```
