---
title: "GARCH Estimation and Filtered Historical Simulation"
author: "Tyler J. Brough"
date: "March 24, 2018"
output:
  pdf_document: default
  html_document: default
---

## Estimating GARCH Models for Filtered Historical Simulation

First we will read in the data and calculate log-returns, the mean return and squared deviations in returns. 


```{r, warning = FALSE}
raw <- read.csv("../data/SP500.csv", header = T)
raw$Date <- as.Date(as.character(raw$Date), format = "%d-%b-%y")
ret <- diff(log(raw$SP500))
raw$Ret <- c(0, ret)
rbar <- mean(ret)
devs <- (ret - rbar)^2
```


### Estimating the GARCH(1,1) Model

We will use the `garch` function in the `tseries` package, and fit a $GARCH(1,1)$ model.

```{r, results = "hide", warning= FALSE}
library(tseries)
fit <- garch(raw$Ret, order = c(1,1))
theta <- coef(fit)
w <- theta[1]
a <- theta[2]
b <- theta[3]
```

There are some helpful methods belonging to the `garch` object in the `tseries` package. We will use the `residuals` method to get the residuals from the $GARCH(1,1)$ model that we will use in the FHS. 

```{r}
resids <- residuals(fit)
hist(resids, breaks = 50)
```

We can then also get the conditional volatilities. 


```{r}
vol.hat <- raw$Ret / resids
hist(vol.hat, breaks = 50)
```

We can now set up the *Filtered Historical Simulation*. Let's simulate one 10-day return path as follows:

```{r, warnings = FALSE}
n <- length(raw$Ret)
r2 <- raw$Ret
s2 <- vol.hat * vol.hat
r0 <- r2[n]
s0 <- s2[n]
s1 <- w + a * r0 + b * s0
z <- sample(resids, size = 10, replace = T)
r <- rep(0, 10)
r[1] <- z[1] * sqrt(s1)
st <- s1

for(t in 2:10)
{
  st <- w + a * (r[t-1] * r[t-1]) + b * st
  r[t] <- z[t] * sqrt(st)
}

r10 <- sum(r)
print(r10)
```
