---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

## Filterd Historical Simulation from a GARCH(1,1) Model

First setup the data, etc.

```{r}
require(rugarch)
## Open the data file
setwd("C:\\Users\\brough\\USU\\Teaching\\Spring2018\\Jike\\Fin6320-Jike")
raw.data <- read.csv("SP500.csv", header = T)
```

The next step is to calculate log-returns. 


```{r}
## Calculate log-returns
ret <- diff(log(raw.data$SP500))
```

Then we must obtain parameter estimates via numerical MLE. We do this using the `rugarch` package. 

```{r}
## Use ugarchfit to estimate the parameters via numerical MLE
sp <- ugarchspec()
fit <- ugarchfit(ret, spec = sp)
w <- coef(fit)[[4]]
a <- coef(fit)[[5]]
b <- coef(fit)[[6]]
c(w, a, b)
```

Now that we have the estimated parameter values, we can set up the FHS. First we need the initial conditions ($r_{0}$, and $\sigma_{0}$).

```{r}
n <- length(ret)
r.init <- ret[n]
v.cond <- as.numeric(fitted(fit))
v.init <- v.cond[n]
c(r.init, v.init)
```

Before we can start the FHS algorithm, we need historical residuals. 

```{r}
resid <- ret / v.cond
tail(resid)
```


Now for the Filtered Historical Simulation algorithm.

```{r}
reps <- 10
steps <- 10
paths <- matrix(0, nrow = reps, ncol = steps+1)
paths[,1] <- r.init

for(i in 1:reps)
{
  z <- sample(resid, size = steps, replace=T)
  sigma <- v.init
  for(j in 2:(steps+1))
  {
    sigma <- sqrt(w + a * (paths[i,j-1] * paths[i,j-1]) + b * (sigma * sigma))
    paths[i,j] <- z[j-1] * sigma
  }
}

ret.pred <- apply(paths[,2:(steps+1)], 1, sum)
```